---
title: 'HW 2: Visualization, Base R'
author:
  name: "HyunYoung Cho"
  email: "chohy@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: default
  pdf_document: default
subtitle: Data 180, Professor Kennedy
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
#Frequency table
table(Cars93$Type)/nrow(Cars93)

#Relative frequency table
round(table(Cars93$Type)/nrow(Cars93),3)
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
barplot(table(Cars93$Type) ,ylab="Frequency" ,cex.lab=1.2 ,col=c(2,3,4) ,main="Car Type")

```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie(table(Cars93$Type) ,main="Car Type" ,col=c(2,3,4) ,cex=1.2)
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.
```{r}
#The mode of the variable 'Type' is Midsize. 
```

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hist((Cars93$Horsepower),cex.lab=1.2,cex.axis=1.2,col="lightblue", xlab="Horsepower", main="Horsepower")
box()
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hist((Cars93$Horsepower),breaks=seq(50,300,25),col="lightblue",xlab="Horsepower",cex.lab=1.2,cex.axis=1.2, main="Horsepower")
box()
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
```{r}
#The histogram of Horsepower with class setting breaks tells R we would like 10 classes. A good rule of thumb is when the number of classes is sqrt(n). The number of classes that follows the good rule of thumb for the histogram of Horsepower is sqrt(93)=9.6. Therefore, the histogram with breaks has the closer number of classes(=10) to the rule of thumb(=sqrt(93)=9.6), the histogram with class breaks follows the rule of thumb for the suggested number of classes. 
```


## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
```{r}
class(Cars93$Passengers)
hist((Cars93$Passengers),cex.lab=1.2,cex.axis=1.2,col="lightblue",xlab="Passengers", main="Passengers")

#The type of variable 'Passengers' is integer, which is numerical and discrete. The histogram of Passengers has x-value(number of passengers) between 3 and 4. However, since the type of variable 'Passengers' is discrete, the x-value(number of passengers) cannot be between 3 and 4. Instead of histogram, a bar graph of 'Passengers' would work better with discrete x-value(number of passengers).
```

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
barplot(table(Cars93$Passengers),ylab="Frequency",cex.lab=1.2,col=c(2,3,4),main="Passengers")
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
hist((Cars93$Length),cex.lab=1.2,cex.axis=1.2,col="lightblue",xlab="Length", main="Length")
box()
```

b) What is the shape of the distribution?
```{r}
#The shape of the distribution of 'Length' histogram is Symmetric and Unimodal.
```

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
```{r}
mean(Cars93$Length)
median(Cars93$Length)
#Yes, since the mean is approximately equal to median, the distribution of variable 'Length' is Symmetric and Unimodal. 
```

You are done! 🏁 Don't forget to commit and push your .Rmd file to your Github repository before the due date.


